Beyesian Inference for Decision Making under COVID-19: A case study

### 2020-05-29 out/paper.Rmd
Draft seccióon V

Section **I** begins by providing the reader a concrete outlook of the problem and the data at hando to tackle it. Section **II** translates the problem into its concrete mathematical and statistical representation and comments on the frailty of classical solution. Section **III** goes about the same formulation under a simple *by-hand* bayesian alternative while section **IV** proposes a more sophisticated solution and cautions the reader on the first asumptions. Finally section **V** is a brief utterance on the importance of providing a well grounded inference for decision making proceses.

@@

In any context not involving such distressfull times everything would work fine, in fact our Fisherian conclusion would be that with 95% confidence level $\theta$ would lie within the **INSERTAR INTERVALO** [^2]. However things start getting funny when incorporating unobserved thesis on the probable behaiviour of $\theta_{COVID-19}$. Lets explore some alternatives:

[^2]: This interval lenght corresponds to the *predictive interval lenght* given by $\frac{X_{n+1} - \overline{X}}{s_n \sqrt{1 + \frac{1}{n}}} \sim T^{n-1}$

@@

Nonetheless many may be still be troubled by the known variance assumption, which is in fact totally reasonable. Relaxing this premise is just a matter of a litter more mathematical intricacy [^7]. For a more realistic scenario we choose the Normal-InverseGamma family where the predictive distribution for $\hat{\theta}_{COVID-19}$ is: $$f(X_{na+1} | X_{(\underline{n})}) \sim  t_{2 \alpha_n}(x | \mu_n, \frac{\beta_n( \kappa_n + 1)}{\alpha_n \kappa_n})$$
Where:

@@

Additionally the posterior ma

[^7]: All normal conjugacy cases are weel documented in **Insertar cita**

>>>>

Section **I** begins by providing the reader a concrete outlook of the problem and the data at hando to tackle it. Section **II** translates the problem into its concrete mathematical and statistical representation and comments on the frailty of classical solution. Section **III** goes about the same formulation under a simple *by-hand* bayesian alternative and introduces the concept ok likelihood resilience, while section **IV** proposes a simpler solution using an alternative conjugate family. Finally section **V** is a brief utterance on the importance of providing a well grounded inference for decision making proceses.

@@

In any context not involving such distressfull times everything would work fine, in fact our Fisherian conclusion would be that with 95% confidence level $\theta$ would lie within the **INSERTAR INTERVALO** [^2]. However things start getting funny when incorporating unobserved hypothesis on the probable behaiviour of $\theta_{COVID-19}$. Lets explore some alternatives:

[^2]: This interval lenght corresponds to the *predictive interval lenght* given by $\frac{X_{n+1} - \overline{X}}{s_n \sqrt{1 + \frac{1}{n}}} \sim T^{n-1}$. The interval is not symetric because of the truncation.

@@

Nonetheless many may be still be troubled by the known variance assumption, which is in fact totally reasonable. Relaxing this premise is just a matter of a litter more mathematical intricacy [^7]. For a more realistic scenario we choose the Normal-InverseGamma family where the predictive distribution for $X_{n+1}$ is: $$f(X_{na+1} | X_{(\underline{n})}) \sim  t_{2 \alpha_n}(x | \mu_n, \frac{\beta_n( \kappa_n + 1)}{\alpha_n \kappa_n})$$
Where:

@@

[^7]: All normal conjugacy cases are weel documented in **Insertar cita**.

As in the known variance case described some lines earlier, we need to truncate the posterior predictive distribution to adjust for $\theta$ 's parametral space yielding the following 95% credibility interval of **INSERTAR INTERVALO** [^8].

[^8]: Truncation of posterior predictive t-distribution was dode with r package *truncdist* v. 1.0-2.

Figure 2 displays a similar likelihood resilience tests for the posterior predictive distribution.

**FIGURA 2**

**DESCRIPCIÓN FIGURA 2**

Although analytical manipulation become a bit more messy than obtainign the MLE from the classical aproach, it should be obvious by know the benefit of advocating for the later methodology. We gain a coherent inference substructure to include unobserved hypotheis, we get the likelihood resilience insurance for free and in the conjugte case we can analytically interaction between model paramters (which in the mark-down fashion we cannot).

# The Exponential-Gamma system

We finally need to adress the parametral space issue as truncation may introduce some distortions of the interest paramter's behaviour. To account for a more realistic representation of $\Theta$, in which case we could opt to characterize $X_i$ as $Exp(\theta)$ however the MLE will still have problem **a** and **b** as described in section I.

The bayesian alternative is to require the Exponential-Gamma conjugate family. Here $X_i$ is distributed as before and $\theta \sim \Gamma(\theta | \alpha_0, \beta_0)$ and the posterior predictive distribution is given by $$f(X_{n+1}| X_{(\underline{n})}) \sim Pareto(x| \alpha_n, \beta_n)$$
Where

  + $\alpha_n = \alpha_o + n -1$
  + $\beta_n = \beta_0 + n \overline{x}$

Given this second conjugate scheme we account for the actual $\Theta$, in fact we simplify the normal conjugate case and again we acquiere the likelihood resilience insurance. This
